---
title: "Modelling the impact of increasing sources using Bayesian, non-linear (power function) regressions"
author: "Manikya Alister"
format: 
  # html:
  #  embed-resources: true
  docx:
    reference-doc: apa-7-template.docx
editor: visual
message: false
warning: false
code-fold: true
prefer-html: true # so that rendering doesn't break for non-html when using html commands
toc: true
echo: false
---

```{r}
library(tidyverse)
library(posterior)
library(here)
library(ggdist)
library(raincloudplots) 
library(knitr)
library(ggpubr)
library(brms)
#library(gtsummary)
library(flextable)

```

# Experiment 1

## Results

Full data and analysis code can be found at \[link\]

### Non-linear Bayesian Modelling

To evaluate the the how relationship between consensus size and participants' confidence in the claim varied as a function of the independence of consensus, we ran a set of Bayesian non-linear regression models. Specifically, we modeled the relationship between consensus size and confidence using a *power*\footnote{We also assessed exponential and logarithmic functions that have slightly different shapes, but the power funciton appeared to fit the data best, although there was only a small differnce in performance between the different functions (see Supplementary Materials).}function:

$$
(start + increase) - increase \times n^{-rate}
$$

Where *start* is a free parameter indicating the initial point of confidence at baseline: confidence when there is one person agreeing and another disagreeing (no consensus). *Increase* is a free parameter indicating the maximum amount that confidence will grow as a function of consensus size, *n*. *Increase* + *Start* therefore represents the asymptote of confidence. *Rate* is a free parameter that controls how quickly confidence grows as a function of consensus size. This equation therefore assumes that confidence increases as a function of consensus size up until a certain point.

We modeled the effect of consensus on each parameter as a fixed effect, where "consensus" refers to the consensus independence condition:

$$
Start \sim 1 + consensus
$$

$$
Increase \sim 1 + consensus
$$

$$
Rate \sim 1 + consensus
$$

This modelling approach has several advantages. First, the power function we use corresponds well to existing theoretical models that have attempted to describe the relationship between conformity and social group size (Bond, 2005; Latané, 1981).

Second, this approach has unique predictive value, allowing us to estimate how increasing the number of sources will increase confidence, even if we do not have direct empirical measurements for a given source number (i.e., our experiment only tests consensus sizes of 4, 7, and 10. This can therefore allow us to do things like predict when adding new sources will stop increasing confidence as a function of different types of consensus manipulations (e.g., independence), with minimal data.

Third, and perhaps most importantly, the parameters in the model are easily interpretable, neatly corresponding to various mechanisms that could be driving different consensus effects. If there are differences in how independence influences the consensus size X confidence relationship, our model allows us to infer the mechanisms that are driving these differences. Specifically, we can infer whether source independence affects the *rate* at which confidence increases with multiple sources, whether it affects the maximum amount that confidence will *increase* as a function of consensus size, or both.

Last, our Bayesian approach is advantageous as it allows us to quantify the posterior probability of different parameter values, rather than quantify the likelihood of a result as extreme as ours assuming the null hypothesis is true, as is the case with *p* values derived from frequentest methods. This advantage is derived from the fact that a Bayesian approach yields full distributions of parameter estimates, which frequentest analyses do not. This distribution also permits intuitive quantification of the uncertainty of our estimates using *credible intervals.* In our analyses, we examine 95% credible intervals, which tell us the range in which, given the data and prior, there is a 95% probability that the parameter falls.

The models were implemented using the R package *brms* (Buerkner 2017; 2018) version 2.23.0, which is a package that facilitates Bayesian modelling by interfacing with the probabilistic programming language Stan (Stan Development Team, 2020; n.d). The analyses were run using R version 4.5.1 (2025-06-13). See the supplementary materials and analysis code for the full implementation of the different models and their priors.

#### Model Fit

As shown in Figure X, our full model, which assumes that all parameters (*start*, *rate*, and *increase)* vary across consensus conditions, provided a good fit of the experimental data. Importantly, it captured the empirical trend whereby confidence increased quickly with initial increases in consensus size before reaching an asymptote. Figure X also suggests that there is a clear distinction between the trajectories of the independent and dependent consensus conditions. In the next section we attempt to uncover what may be driving these differences.

```{r}
source(here("analyses/plot-fits.R")) # source function that plots the model fit 

plotFit(exp = 1,
        title = "Experiment 1",
        subtitle = NULL)
```

#### Model Comparison

In Figure X, our modelling suggests that there is a clear distinction between the trajectories of the independent and dependent consensus conditions respectively. However, it is not clear what is driving these differences. Specifically, it is unclear whether independence results in *faster* growth in confidence (i.e., is driven by the *rate* parameter), results in higher maximum increases in confidence (i.e., is driven by the *increase* parameter) or is driven by both mechanisms.

To test these possibilities, we conducted a model comparison whereby we ran models that assumed that different combinations of parameters varied across consensus conditions. For example, in the *increase* model, *increase* was the only parameter that could vary across consensus conditions, meaning it was the only parameter that could explain any differences in that model. Similarly, the *increase + rate* model assumed that both *increase* and *rate* varied across conditions. We also had a *full* model that assumed all parameters varied across conditions, a *null* model, that assumed no changes in any parameter across conditions, and all other possible combinations (see Table X). We then compared model performance using Leave One Out Cross Validation (LOOIC), which balances model complexity with model fit. This approach therefore allowed us to examine 1) whethere any differences between conditions were statistically reliable, and 2) which parameters were driving these differences.

The results of the model comparison are in @tbl-m-comp-e1. It shows that the best performing model in Experiment 1 according to LOOIC was the *increase* model, that assumes only the increase parameter varies across conditions. This result suggests that 1) there are statistically credible differences in the relationship between consensus size and confidence between independent and dependent consensuses and 2) that independence results in a larger maximum increase in confidence, but not a *faster* increase in confidence relative to the dependent consensus. It is also a good sanity check that differences in conditions are not driven by differences in their baseline confidence (*start*) since there is no consensus independence manipulation in the baseline 1:1 trial.

```{r}
#| label: tbl-m-comp-e1
#| tbl-cap: "Model comparison using LOOIC (lower LOOIC = better performance). "

load(here("analyses/output/looic-model-comparison-e1.Rdata"))
load(here("analyses/output/weighted-looic-model-comparison-e1.Rdata"))
model_full_names <- c("Null", "Rate", "Increase", "Start", "Start + Increase", "Start + Rate", "Rate + Increase", "Full Model")

modelComparisonTable= function(model_looic, model_full_names){
  model_comparison <- model_looic %>%
  mutate(
        Rank = rank(LOOIC),
        LOOIC = as.character(round(LOOIC, 2)), # as chatacter to stop it from ading a "," in LOOIC value
        SE = round(SE, 2),
        Model = model_full_names
        ) %>%
  select(-experiment, -model) %>%
relocate(Model) 

autofit(flextable(model_comparison))
}

modelComparisonTable(model_looic, model_full_names)

```

@tbl-params-e1 shows the parameter estimates of the best performing model (the *Increase*) model, and the 95% credible intervals of those estimates. The estimates predicted that confidence grows from the scale mid point (50), reaching a maximum (asymptote) confidence of 73 for the dependent consensus and 83.99 for the independent consensus. As mentioned earlier, this highlights a benefit of our modelling approach, since the maximum estimated confidence values (the point at which confidence will no longer increase as a function of the consensus size, calculated as *Start* + *Increase*) was outside of the observed measured confidence (see Figure X). The differences in *Increase* across conditions appeared to be statistically reliable, with very little overlap between the 95% credible intervals for the Increase parameter in the Independent and Dependent consensus conditions respectively.

```{r}
#| label: tbl-params-e1
#| tbl-cap: "Parameter estimates of the best performing model in Experument 1: the *Increase* model, in which the Increase parameter is the only parameter that can vary across consensus independence conditions. This change across conditions is operationalised by estimating Increase for the baseline condition (the *Dependent* condition), then estimating the *change* in that paremter in the independent condition. "

load(here("analyses/output/fe-brms-exp1-power-fit-increase.Rdata"))
param_names <- c("Start",
                 "Increase_{Dependent}",
                 "Increase_{Independent}",
                 "Rate")


extractEstimates = function(fit, param_names, target_param = "increase") {
  summ_fit <- summary(fit)
  estimates <- summ_fit$fixed 
  
  exp_estimates <- round(exp(estimates),2)

  backup_estimates <- summ_fit$fixed # preserve unnexponentiated estimates for calculating the condition effect 
  params <- rownames(estimates)
  effect_rows <- grepl("consensus", params)
  effect_params <- params[effect_rows]
  baseline_intercept <- paste0(target_param, "_Intercept")
  for (i in 1:length(effect_params)) {
    param <- effect_params[i]
    # adjust the estimate so it is on original scale and reflects independence effect
    exp_estimates[param, "Estimate"] <- exp(estimates[param, "Estimate"] + estimates[baseline_intercept, "Estimate"])
    exp_estimates[param, "l-95% CI"] <- exp(estimates[param, "Estimate"] + estimates[baseline_intercept, "l-95% CI"])
    exp_estimates[param, "u-95% CI"] <- exp(estimates[param, "Estimate"] + estimates[baseline_intercept, "u-95% CI"])
  }
  
  exp_estimates <- round(exp_estimates,2)
  
  clean_estimates <- exp_estimates %>%
    mutate(Parameters = param_names) %>%
    select(-Rhat, -Bulk_ESS, -Tail_ESS, -Est.Error)  %>%
    relocate(Parameters) %>%
    remove_rownames() 
  
  autofit(flextable(clean_estimates))
  
}

extractEstimates(fit, param_names)

```

# Experiment 2

## Results

### Non-linear Bayesian Modelling

We conducted the same non-linear Bayesian regression modelling as in Experiment 1.

#### Model Fit

The model fit similarly well as in Experiment 1, if not slightly better (see Figure X).

```{r}
plotFit(exp = 2,
        title = "Experiment 2",
        subtitle = NULL)
```

#### Model Comparison

We performed the same model comparison (with the same models) as in Experiment 1. The qualitative results were basically identical, with the *Increase* *model* performing the best, followed by the *Rate +* *Increase* model.

```{r}
#| label: tbl-m-comp-e2
#| tbl-cap: "Model comparison using LOOIC (lower LOOIC = better performance). "

load(here("analyses/output/looic-model-comparison-e2.Rdata"))

modelComparisonTable(model_looic, model_full_names)
```

As before, the parameters of the best performing (*increase*) model are in @tbl-m-comp-e2. The estimates in Experiment 2 were very similar to Experiment 1, predicting a maximum confidence (asymptote) of 72.95 for the dependent consensus condition, and 84.77 for the independent consensus condition. Once again, these results appeared to be statistically reliable, with very little overlap between the 95% credible intervals for the Increase parameter in the Independent and Dependent consensus conditions respectively.

```{r}
#| label: tbl-params-e2
#| tbl-cap: "Parameter estimates of the best performing model in Experument 2: the *Increase* model, in which the Increase parameter is the only parameter that can vary across consensus independence conditions. This change across conditions is operationalised by estimating Increase for the baseline condition (the *Dependent* condition), then estimating the *change* in that paremter in the independent condition. "

load(here("analyses/output/fe-brms-exp2-power-fit-increase.Rdata"))
param_names <- c("Start",
                 "Increase_{Dependent}",
                 "Increase_{Independent}",
                 "Rate")



extractEstimates(fit, param_names)
```

# Experiment 3

## Results

### Non-linear Bayesian Modelling

We once again implemented the same non-linear Bayesian regression models as in Experiments 1 and 2. Ideally, in Experiment 3 we would have run a hierarchical model that estimated each parameter both at the group level, and at the level of each individual claim. However, we had insufficient observations for such an approach, which would have therefore resulted in convergence issues unless stronger priors were implemented, in which case the estimates would have been heavily influenced by the priors rather than reflecting any empirical observations.

#### Model Fit

As per Figure X, the model provides a good fit for all three consensus independence conditions.

```{r}
plotFit(exp = 3,
        title = "Experiment 2",
        subtitle = NULL)
```

#### Model Comparison

We conducted the same model comparison (with the same models) as experiments 1 and 2 (see @tbl-m-comp-e3). Once again, the *Increase* model performed best, suggesting that differences across all three conditions were likely to be driven by differences in the maximum growth in confidence as a function of increased consensus size. However, it is worth noting that in Experiment 3, the next best model was the *Rate* model, and that the differences between the *Rate* and *Increase* model were very small. It is therefore possible that changes across conditions were driven by both *Rate* and *Increase,* but the model comparison did not favor the *Rate + Increase* model due to insufficient power (LOOIC penalizes complex models more when *n* is smaller; Vehtari et al., 2017).

```{r}
#| label: tbl-m-comp-e3
#| tbl-cap: "Model comparison using LOOIC (lower LOOIC = better performance). "

load(here("analyses/output/looic-model-comparison-e3.Rdata"))
modelComparisonTable(model_looic, model_full_names)
```

```{r}
#| label: tbl-params-e3
#| tbl-cap: "Parameter estimates of the best performing model in Experument 3: the *Increase* model, in which the Increase parameter is the only parameter that can vary across consensus independence conditions. This change across conditions is operationalised by estimating Increase for the baseline condition (the *Dependent* condition), then estimating the *change* in that paremter in the independent condition. "

load(here("analyses/output/fe-brms-exp3-power-fit-increase.Rdata"))
param_names <- c("Start",
                "Increase_{Dependent}",
                 "Increase_{Dependent Source}",
                 "Increase_{Independent}",
                 "Rate")
extractEstimates(fit, param_names) # PROBLEM: Exp is obscuring legit negative fixed effects. 
# Also, should I be exponentiating the errors??? 
# Soome follow up on Exp 3 might be important so that we know if increase is best for ALL conditions 
```

These results suggest that increase is driving differences between all three consensus conditions in Experiment 3. However, it's possible that there are divergent mechanisms driving the differences between Dependent Source vs. Dependent and Dependent vs. Independent. Our original model comparison would not have been able to pick up on those differences. For example, it could be that *Increase* only explains the difference between independent vs. dependent, not dependent vs. dependent source.

To test this possibility, we ran an identical model comparisons comparing independent vs. dependent conditions and dependent vs. dependent_source in Experiment 3 respectively. @tbl-m-comp-e3-rm-dep-source shows the model comparison for independent vs. dependent. As in all of the previous experiments, the *Increase* model performed best. Consistent with Experiments 1 and 2, the Rate + Increase model was second best, suggesting that we could qualitatively replicate the trends of the independent vs. dependent consensus when aggregating across all three claims.

```{r}
#| label: tbl-m-comp-e3-rm-dep-source
#| tbl-cap: "Experiment 3 Model comparison using LOOIC (lower LOOIC = better performance), comparing Independent vs. Dependent (excluding Dependent Source)"
load(here("analyses/output/looic-model-comparison-e3-rm-dependent_source.Rdata"))
modelComparisonTable(model_looic_rm_dependent_source, model_full_names)
```

@tbl-m-comp-e3-rm-ind shows the model comparison comparing the dependent vs. dependent source conditions. This comparison shows results that are quite different to the independent vs. dependent comparison, and all previous experiments. In this comparison, the *start* model performed best, suggesting that the most likely mechanism driving any differences between these conditions was differences in the baseline condition. Given that the baseline conditions were identical in all conditions (1 source arguing for and 1 arguing against) this difference is most likely spurious. Importantly, one interpretation of this result is that the effect of increasing consensus size on confidence is *functionally equivalent* for dependent source versus the standard dependent consensus condition.

```{r}
#| label: tbl-m-comp-e3-rm-ind
#| tbl-cap: "Experiment 3 Model comparison using LOOIC (lower LOOIC = better performance), comparing Dependent vs. Dependent Source (excluding Independent)"
load(here("analyses/output/looic-model-comparison-e3-rm-independent.Rdata"))
modelComparisonTable(model_looic_rm_independent, model_full_names)
```


# References

-   Bond, R. (2005). Group Size and Conformity. *Group Processes & Intergroup Relations*, *8*(4), 331–354. <https://doi.org/10.1177/1368430205056464>

-   Latané, B. (1981). The psychology of social impact. *American Psychologist*, *36*(4), 343–356. <https://doi.org/10.1037/0003-066X.36.4.343>

-   Paul-Christian Buerkner (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. *Journal of Statistical Software*, 80(1), 1-28. `doi:10.18637/jss.v080.i01`

-   Paul-Christian Buerkner (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. *The R Journal*. 10(1), 395–411. `doi:10.32614/RJ-2018-017`

-   The Stan Development Team (n.d). *Stan Modeling Language User's Guide and Reference Manual*. <https://mc-stan.org/users/documentation/>.

-   Stan Development Team (2020). RStan: the R interface to Stan. R package version 2.21.2. <https://mc-stan.org/>

-   Vehtari, A., Gelman, A., & Gabry, J. (2017). Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC. *Statistics and Computing*, *27*(5), 1413–1432. <https://doi.org/10.1007/s11222-016-9696-4>

# Supplementary Materials

I have just thrown in a few extra analyses/information that might be useful to include in the supplementary materials (you might also think they belong in the main manuscript, too). They haven't all been linked in text though.

# Comparing prior belief vs. baseline

A relevant question for relevant for our modelling approach is how to deal with participants' reported prior beliefs compared to the baseline condition, where there is 1 person agreeing with the claim, and 1 person disagreeing. In theory, it would be nice to control for people's prior beliefs in some way. In all of my previous consensus studies we have done this, in order to account for the fact that people who already have strong prior beliefs are less likely to shift them. There are a few problems with this in the current study, though.

1.  There is only one trial per participant, so you cannot control for individual variation in priors in the same way as you normally would (i.e., with a random intercept).
2.  As I will show, the prior and baseline condition are *highly* correlated, meaning that controlling for controlling for prior belief as a fixed effect could be problematic, and artificlally reduce the predictive power of the model and result in issues of multicollinearity.

## Does confidence change from prior to baseline?

If we can treat prior and baseline as essentially the same, this will vastly simplify our modelling, as we can treat the baseline condition as people's priors. First, let's check the correlation

```{r}
experiments = 1:3

source(here("analyses/prior-baseline-comparison.R"))

all_corr_plots <- vector("list", length = length(experiments))

for (i in 1:length(experiments)){
  all_corr_plots[[i]] <- checkPriorBaselineCorr(i)
}


combined_corr_plot <- ggarrange(plotlist = all_corr_plots)
combined_corr_plot
```

These plots show very high correlations between someone's prior belief and their confidence at baseline. Now lets compare the means of the groups and their distributions.

```{r}
all_group_plots <- vector("list", length = length(experiments))

for (i in 1:length(experiments)){
  exp <- experiments[i]
  all_group_plots[[i]] <- checkPriorBaselineGroups(exp)
}


combined_group_plot <- ggarrange(plotlist = all_group_plots)
combined_group_plot
```

These plots show that scores are pretty much the same from prior to baseline. Although some participants do seem to change their rating, this appears to be the vast minority, and at the group level the distributions look similar. The means are also not significantly different from each other.

Something else that's interesting is that the baseline condition seems to bring the distribution more towards the middle, somewhat nullifying the effect of priors (i.e., the baseline distribution is more symmetrical around 0).

Taken together, I would argue that it's safe to ignore prior beliefs in the modelling and instead focus only on how beliefs change from baseline.

### Weighted Model Comparison

The model comparisons in the paper are interpreted using raw LOOIC values. However, many people might find it difficult to compare the scale of  relative performance for each model. In othwe words, it's unclear how big a difference of, for example, 1 LOOIC is. An alternative to this is to convert the raw LOOIC into relative model weights. This therefore represents performance as a kind of "relative probability" of one model compared to the other. We could consider adding these weights to the model comparison tables. 

Model weights experiment 1: 
```{r}
load(here("analyses/output/weighted-looic-model-comparison-e1.Rdata"))
weightTable = function(d_weighted_looic){
  weight_table <- d_weighted_looic %>%
  mutate(Weight = round(Weight,2)) %>%
  relocate(Model)
autofit(flextable(weight_table))
}
weightTable(d_weighted_looic)
```
Model weights experiment 2: 
```{r}
load(here("analyses/output/weighted-looic-model-comparison-e2.Rdata"))
weightTable(d_weighted_looic)
```
Model weights experiment 3: 
```{r}
load(here("analyses/output/weighted-looic-model-comparison-e3.Rdata"))
weightTable(d_weighted_looic)

```

Something else that we can do with these weights is create model groupings. So we can combine all of the models that assume *Increase* changes across conditions (Increase, Increase+start, Increase+rate, Full), and compare that to models that allow other parameters to change across conditions. This can give us a measure of the relative imporatance of each parameter: 

```{r}
load(here("analyses/output/combined-weighted-group-plot-exp-1-2-3.Rdata"))
combined_weighted_group_plot
```


## Models

In this section, we show the precise formalization of the different models that we ran in our model comparison, as they were implemented in BRMS. Note that in our models, all parameters were exponentiated so that they are constrained to be positive. Full model code can be found at \[link to model code\].

All models assumed a power function relationship between consensus size and confidence, as follows:

```         
  confidence ~ (exp(start) + exp(increase)) - exp(increase) * (n_sources^(-exp(rate)))
```

What changed between models was how the fixed effects were defined. In other words, does the model assume the parameters vary as a function consensus condition?

### *None* model: assumes neither *start, increase,* nor *rate* vary as a function of consensus

```         
      start ~ 1,
      increase ~ 1, 
      rate ~ 1, 
```

#### *Start* model: assumes that only the *start* parameter varies as a function of consensus.

```         
 start ~ 1 + consensus
 increase ~ 1
  rate ~ 1
```

#### *Increase* model: assumes that only the *increase* parameter varies as a function of consensus.

```         
start ~ 1
increase ~ 1 + consensus
rate ~ 1
```

#### *Rate* model: assumes that only the *rate* parameter varies as a function of consensus.

```         
start ~ 1
increase ~ 1
rate ~ 1 + consensus
```

#### *Start + Increase* model: assumes that both the *start* and *increase* parameters vary as a function of consensus.

```         
start ~ 1 + consensus
increase ~ 1 + consensus
rate ~ 1
```

#### *Start + Rate* model: assumes that both the *start* and *rate* parameters vary as a function of consensus.

```         
start ~ 1 + consensus
increase ~ 1
rate ~ 1 + consensus
```

#### *Increase + Rate* model: assumes that both the *increase* and *rate* parameters vary as a function of consensus.

```         
start ~ 1
increase ~ 1 + consensus
rate ~ 1 + consensus
```

#### *Full* model: assuming *start, increase,* and *rate* all vary as a function of consensus condition.

```         
      start ~ 1 + consensus,
      increase ~ 1 + consensus,
      rate ~ 1 + consensus
```

#### 

### Priors

We placed weakly informative priors on all parameters. Intercepts were given log-normal priors reflecting plausible values, while consensus effects (slopes) were centered at zero with moderate variance:

-   **Intercepts**

$$
 Start \sim \mathcal{N}(\log 50, 0.5) 
$$
$$
  Increase \sim \mathcal{N}(\log 25, 0.5) 
$$

$$
  Rate \sim \mathcal{N}(\log 0.5, 0.2) 
$$

-   **Condition Effects (when included)**\

$$
\mathcal{N}(0, 0.5) 
$$
