---
title: "Modelling the impact of increasing sources using Bayesian, non-linear (power function) regressions"
author: "Manikya Alister"
format: 
  # html:
  #  embed-resources: true
  docx:
    reference-doc: apa-7-template.docx
editor: visual
message: false
warning: false
code-fold: true
prefer-html: true # so that rendering doesn't break for non-html when using html commands
toc: true
echo: false
---

```{r}
library(tidyverse)
library(posterior)
library(here)
library(ggdist)
library(raincloudplots) 
library(knitr)
library(ggpubr)
library(brms)
#library(gtsummary)
library(flextable)

```

# Experiment 1

## Results

Full data and analysis code can be found at \[link\]

### Non-linear Bayesian Modelling

To evaluate how the relationship between consensus size and participants' confidence in the claim varied as a function of the independence of consensus, we ran a set of Bayesian non-linear regression models. Specifically, we modeled the relationship between consensus size and confidence using a *power*\footnote{We also assessed exponential and logarithmic functions that have slightly different shapes, but the power funciton appeared to fit the data best, although there was only a small differnce in performance between the different functions (see Supplementary Materials).}function:

$$
(start + increase) - increase \times n^{-rate}
$$

Where *start* is a free parameter indicating the initial point of confidence at baseline: confidence when there is one person agreeing and another disagreeing (no consensus). *Increase* is a free parameter indicating the maximum amount that confidence will grow as a function of consensus size, *n*. *Increase* + *Start* therefore represents the asymptote of confidence. *Rate* is a free parameter that controls how quickly confidence grows as a function of consensus size. This equation therefore assumes that confidence increases as a function of consensus size up until a certain point or *asymptote* (*start* + *Increase*).

We modeled the effect of consensus on each parameter as a fixed effect, where "consensus" refers to the consensus independence condition:

$$
Start \sim 1 + consensus
$$

$$
Increase \sim 1 + consensus
$$

$$
Rate \sim 1 + consensus
$$

This modelling approach has several advantages. First, the power function we use corresponds well to existing theoretical models that have attempted to describe the relationship between conformity and social group size (Bond, 2005; Latané, 1981).

Second, this approach has unique predictive value, allowing us to estimate how increasing the number of sources will increase confidence, even if we do not have direct empirical measurements for a given source number (i.e., our experiment only tests consensus sizes of 4, 7, and 10. This can therefore allow us to do things like predict when adding new sources will stop increasing confidence as a function of different types of consensus manipulations (e.g., independence), with minimal data.

Third, and perhaps most importantly, the parameters in the model are easily interpretable, neatly corresponding to various mechanisms that could be driving different consensus effects. If there are differences in how independence influences the consensus size X confidence relationship, our model allows us to infer the mechanisms that are driving these differences. Specifically, we can infer whether source independence affects the *rate* at which confidence increases with multiple sources, whether it affects the maximum amount that confidence will *increase* as a function of consensus size, or both.

Last, our Bayesian approach is advantageous as it allows us to quantify the posterior probability of different parameter values, rather than quantify the likelihood of a result as extreme as ours assuming the null hypothesis is true, as is the case with *p* values derived from frequentest methods. This advantage is derived from the fact that a Bayesian approach yields full distributions of parameter estimates, which frequentest analyses do not. This distribution also permits intuitive quantification of the uncertainty of our estimates using *credible intervals.* In our analyses, we examine 95% credible intervals, which tell us the range in which, given the data and prior, there is a 95% probability that the parameter falls.

The models were implemented using the R package *brms* (Buerkner 2017; 2018) version 2.23.0, which is a package that facilitates Bayesian modelling by interfacing with the probabilistic programming language Stan (Stan Development Team, 2020; n.d). The analyses were run using R version 4.5.1 (2025-06-13). See the supplementary materials and analysis code for the full implementation of the different models and their priors.

#### Model Fit

As shown in Figure X, our full model, which assumes that all parameters (*start*, *rate*, and *increase)* vary across consensus conditions, provided a good fit of the experimental data. Importantly, it captured the empirical trend whereby confidence increased quickly with initial increases in consensus size before reaching an asymptote. Figure X also suggests that there is a clear distinction between the trajectories of the independent and dependent consensus conditions. In the next section we attempt to uncover what may be driving these differences.

```{r}
source(here("analyses/plot-fits.R")) # source function that plots the model fit 

plotFit(exp = "1-combined",
        title = "Experiment 1",
        subtitle = NULL,
        model = "full")
```

We can also separate by valence, as shown in Figure X. Because the model is averaging over valence, it the fits look worse when faceting across them, but importantly the trajectories of both valence levels look similar.

```{r}
plotFit(exp = "1-combined",
        title = "Experiment 1",
        subtitle = NULL,
        model = "full",
        by_group = "posneg")
```

#### Model Comparison

In Figure X, our modelling suggests that there is a clear distinction between the trajectories of the independent and dependent consensus conditions respectively. However, it is not clear what is driving these differences. Specifically, it is unclear whether independence results in *faster* growth in confidence (i.e., is driven by the *rate* parameter), results in higher maximum increases in confidence (i.e., is driven by the *increase* parameter) or is driven by both mechanisms.

To test these possibilities, we conducted a model comparison whereby we ran models that assumed that different combinations of parameters varied across consensus conditions. For example, in the *increase* model, *increase* was the only parameter that could vary across consensus conditions, meaning it was the only parameter that could explain any differences in that model. Similarly, the *increase + rate* model assumed that both *increase* and *rate* varied across conditions. We also had a *full* model that assumed all parameters varied across conditions, a *null* model, that assumed no changes in any parameter across conditions, and all other possible combinations (see Table X). We then compared model performance using Leave One Out Cross Validation (LOOIC), which balances model complexity with model fit. This approach therefore allowed us to examine 1) whethere any differences between conditions were statistically reliable, and 2) which parameters were driving these differences.

The results of the model comparison are in @tbl-m-comp-e1. It shows that the best performing model in Experiment 1 according to LOOIC was the *Rate + Increase* model, that assumes that both the increase and rate parameters vary across consensus conditions. This result suggests that 1) there are statistically credible differences in the relationship between consensus size and confidence between independent and dependent consensuses and 2) that independence results in a larger maximum increase in confidence and potentially also a *faster* increase in confidence relative to the dependent consensus. It is also a good sanity check that differences in conditions are not driven by differences in their baseline confidence (*start*) since there is no consensus independence manipulation in the baseline 1:1 trial.

It is worth pointing out however, that the difference between the best performing *Increase + Start* model is very similar to the *Increase* only model. The Weight column in @tbl-m-comp-e1 shows pseudo-Bayesian model averaging weights, which can be interpreted as approximate posterior probabilities that each model provides the best predictions. The *Rate + Increase* and *Increase* models have similar probabilities, indicating substantial uncertainty about whether it is truly necessary for the rate parameter to differ between consensus conditions. Additionally, comparing the ΔELPD, which represents the difference in expected log point wise predictive density from the best model, reveals that the Increase model performs nearly identically to Rate + Increase (ΔELPD = -0.02 ± 0.12). The small magnitude of this difference relative to its standard error suggests the two models are effectively equivalent in predictive performance. Together, these results provide strong evidence that consensus independence increases the maximum confidence gain (increase parameter), with more uncertain evidence for an additional effect on the rate of confidence accumulation.

```{r}
#| label: tbl-m-comp-e1
#| tbl-cap: "Model comparison statistics for Experiment 1. Models are ranked by Leave-One-Out Information Criterion (LOOIC), with lower values indicating better predictive performance. SE = standard error of LOOIC. ΔELPD = difference in expected log pointwise predictive density from the best-fitting model, with values closer to zero indicating similar performance. SE(ΔELPD) = standard error of the difference in ELPD. Weight = relative probability of each model relative to the other models, caluclated using pseudo Bayesian model avereging. Models varying only the rate and increase parameters by consensus showed the best predictive performance. "

load(here("analyses/output/looic-model-comparison-e1-combined.Rdata"))
model_full_names <- c(
  "none" = "Null",
  "rate" = "Rate",
  "increase" = "Increase",
  "start" = "Start",
  "start_increase" = "Start + Increase",
  "start_rate"  ="Start + Rate",
  "rate_increase" = "Rate + Increase",
  "full"="Full Model"
)

modelComparisonTable <- function(model_looic, model_full_names) {
  raw_looic <- model_looic[[2]]
  weights <- loo_model_weights(model_looic[[3]],method = "pseudobma")
  formal_compare <- model_looic[[1]]
  
  # Convert loo_compare output to data frame with model names
  # loo_compare orders models from best (top) to worst (bottom)
  formal_compare_df <- as.data.frame(formal_compare)
  formal_compare_df$model <- rownames(formal_compare_df)
  
  # Join everything together
  model_comparison <- raw_looic %>%
    left_join(
      formal_compare_df %>% 
        select(model, elpd_diff, se_diff),
      by = "model"
    ) %>%
    mutate(
      Rank = rank(LOOIC),
      LOOIC = as.character(round(LOOIC, 2)),
      SE = round(SE, 2),
      Model = model_full_names[model],
      Weight = round(as.numeric(weights[model]),2),
      `ΔELPD` = round(elpd_diff, 2),
      `SE(ΔELPD)` = round(se_diff, 2)
    ) %>%
    arrange(Rank) %>%  # Sort from best to worst
    select(Model, Rank, LOOIC, SE, `ΔELPD`, `SE(ΔELPD)`, Weight)
  
  flextable::autofit(flextable::flextable(model_comparison))
}

modelComparisonTable(model_looic, model_full_names)
```

@tbl-params-e1 shows the parameter estimates of the best performing model (the Rate + Increase model), and the 95% credible intervals of those estimates. The estimates predicted that confidence grows from approximately 52.64 at baseline (Start), reaching a maximum (asymptote) confidence of 74.76 for the dependent consensus (52.64 + 22.12) and 81.76 for the independent consensus (52.64 + 30.02). As mentioned earlier, this highlights a benefit of our modelling approach, since the maximum estimated confidence values (the point at which confidence will no longer increase as a function of the consensus size, calculated as Start + Increase) was outside of the observed measured confidence (see Figure X). The differences in Increase across conditions appeared to be statistically reliable, with no overlap between the upper 95% credible interval of Increase\_{Dependent} and the Increase\_{Independent} estimate, and no overlap between the lower 95% credible interval for Increase\_{Independent} and the Increase\_{Dependent} estimate.The Rate parameter, which governs how quickly confidence approaches its asymptote, showed similar estimates across conditions, with the median estimates of the Dependent and Independent conditions overlapping with the others' credible intervals, suggesting the rate of confidence accumulation was less statistically reliable.

```{r}
#| label: tbl-params-e1
#| tbl-cap: "Parameter estimates of the best performing model in Experument 1: the *Increase* model, in which the Increase parameter is the only parameter that can vary across consensus independence conditions. This change across conditions is operationalised by estimating Increase for the baseline condition (the *Dependent* condition), then estimating the *change* in that paremter in the independent condition. "

load(here("analyses/output/fe-brms-exp-1-combined-power-fit-rate_increase.Rdata"))
param_names <- c("Start",
                 "Increase_{Dependent}",
                 "Increase_{Independent}",
                 "Rate_{Dependent}",
                 "Rate_{Independent}")


extractEstimates = function(fit, param_names) {
  summ_fit <- summary(fit)
  estimates <- summ_fit$fixed 
  
  exp_estimates <- round(exp(estimates), 2)
  
  params <- rownames(estimates)
  effect_rows <- grepl("consensus", params)
  effect_params <- params[effect_rows]
  
  draws <- as_draws_df(fit)
  
  for (i in 1:length(effect_params)) {
    param <- effect_params[i]
    base_param <- sub("_.*", "", param)
    baseline_intercept <- paste0(base_param, "_Intercept")
    
    draw_col_slope     <- paste0("b_", param)
    draw_col_intercept <- paste0("b_", baseline_intercept)
    
    combined_draws <- exp(draws[[draw_col_intercept]] + draws[[draw_col_slope]])
    
    exp_estimates[param, "Estimate"]   <- round(mean(combined_draws), 2)
    exp_estimates[param, "l-95% CI"]  <- round(quantile(combined_draws, 0.025), 2)
    exp_estimates[param, "u-95% CI"]  <- round(quantile(combined_draws, 0.975), 2)
  }
  
  exp_estimates <- round(exp_estimates, 2)
  
  clean_estimates <- exp_estimates %>%
    mutate(Parameters = param_names) %>%
    select(-Rhat, -Bulk_ESS, -Tail_ESS, -Est.Error) %>%
    relocate(Parameters) %>%
    remove_rownames()
  
  autofit(flextable(clean_estimates))
}

extractEstimates(fit, param_names)

```

# Experiment 2

## Results

### Non-linear Bayesian Modelling

We implemented the same non-linear Bayesian regression models as in Experiment 1. Due to a limited number of observations, we elected to average over claim type rather than control for potential claim type effects through random effects or a hierarchical model. We believe this was a reasonable approach because differences in claim type were not of theoretical interest, and including more complex models that attempted to control for or assess claim type differences would have reduced our power to detect differences in consensus conditions, which was our primary aim. This approach was also justified given that our design was balanced across claim and consensus conditions.

#### Model Fit

As per Figure X, the model provides a good fit for all three consensus independence conditions.

```{r}
plotFit(exp = 2,
        title = "Experiment 2",
        subtitle = NULL,
        model = "full")
```

We can also look at how the overall model performs as a function of each claim type, as shown in Figure X. However it is important to note that the model *averages over claim type* so the fits will inevitably look worse when separating by claim type, since the model has no ability to account for those differences.

```{r}
plotFit(exp = 2,
        title = "Experiment 2",
        subtitle = NULL,
        model = "full",
        by_group = "claim")
```

#### Model Comparison

We conducted the same model comparison (with the same models) as Experiment 1 (see @tbl-m-comp-e2). This time, the *Start + Increase* model performed best, suggesting that differences across all three conditions were likely to be driven by differences in the maximum growth in confidence as a function of increased consensus size, and differences in the starting point. However, similar to Experiment 1, the ratio between ΔELPD and SE(ΔELPD) suggests the second best performing model (the *Full* model) performed equivalently well. 

```{r}
#| label: tbl-m-comp-e2
#| tbl-cap: "Model comparison using LOOIC (lower LOOIC = better performance). "

load(here("analyses/output/looic-model-comparison-e2.Rdata"))
modelComparisonTable(model_looic, model_full_names)
```

```{r}
#|include: FALSE
#| label: tbl-params-e2
#| tbl-cap: "Parameter estimates of the best performing model in Experument 2: the *Start + Increase* model, in which the Increase parameter and Start parameter can vary across consensus independence conditions. This change across conditions is operationalised by estimating Increase and Start for the baseline condition (the *Dependent* condition), then estimating the *change* in that paremter in the independent condition. "

load(here("analyses/output/fe-brms-exp-2-power-fit-start_increase.Rdata"))
param_names <- c("Start_{Dependent}",
                 "Start_{Dependent Source}",
                 "Start_{Independent}",
                "Increase_{Dependent}",
                 "Increase_{Dependent Source}",
                 "Increase_{Independent}",
                 "Rate")
extractEstimates(fit, param_names) 
```

It is possible that there are divergent mechanisms driving the differences between Dependent Source vs. Dependent and Dependent vs. Independent. If this were the case, our original model comparison in @tbl-m-comp-e2 would not have been able to pick up on those differences. For example, it could be that *Increase* only explains the difference between independent vs. dependent, not dependent vs. dependent source.

To test this possibility, we ran an identical model comparisons comparing independent vs. dependent conditions and Dependent vs. Dependent Source in Experiment 3 respectively. @tbl-m-comp-e3-rm-dep-source shows the model comparison for independent vs. dependent. This comparison reveals a similar result to Experiment 1, with the *Rate + Increase* model performing best (26% probability), though the Increase* model performed nearly identically (26% probability, ΔELPD = -0.01 ± 0.14).  This pattern is consistent with Experiment 1, where there was support for both *Increase* and *Rate* driving differences in conditions, with stronger statistical support for *Increase* relative to *Rate.* These results suggest that the mechanism by which independent consensus increases confidence relative to dependent consensus may involve both a higher asymptotic confidence and potentially a faster rate of accumulation.

```{r}
#| label: tbl-m-comp-e3-rm-dep-source
#| tbl-cap: "Experiment 2 Model comparison using LOOIC (lower LOOIC = better performance), comparing Independent vs. Dependent (excluding Dependent Source)"
load(here("analyses/output/looic-model-comparison-e2-rm-dependent_source.Rdata"))
modelComparisonTable(model_looic_rm_dependent_source, model_full_names)
```

```{r}
#| label: tbl-params-e2-rm-dep-source
#| tbl-cap: "Parameter estimates of the best performing model in Experument 2 comparing Independent vs. Dependent Consensus: the *Rate + Increase* model, in which the Increase and Rate parameters can vary across consensus independence conditions."

load(here("analyses/output/fe-brms-exp-2-power-fit-rate_increase-rm-dependent_source.Rdata"))
param_names <- c("Start",
                "Increase_{Dependent}",
                "Increase_{Independent}",
                 "Rate_{Dependent}",
                "Rate_{Inependent}")
extractEstimates(fit, param_names)
```

@tbl-m-comp-e3-rm-ind shows the model comparison for Dependent vs. Dependent Source conditions. This comparison reveals a different pattern from both Experiment 1 and the Independent vs. Dependent comparison. The Start model performed best (24% probability), suggesting that the primary difference between these conditions is in baseline confidence rather than in how confidence grows with consensus size. This is likely due to sampling error, given the baseline ratings were exactly the same in all conditions (i.e., there was no consensus in baseline).

However, all models showed similar predictive performance, with ΔELPD values ranging from -0.18 to -1.97, all well within their standard errors (except for the null model). The model weights are distributed relatively evenly across models, with no single model receiving substantial support. This lack of clear differentiation suggests that the effect of source dependency (whether the same person cites the same source repeatedly vs. different people citing the same source) may be weak or operate through multiple parameters simultaneously. The relatively poor performance of models focusing solely on Increase or Rate parameters (ranks 3 and 2 respectively) contrasts with the Independent vs. Dependent comparison, suggesting that source dependency affects confidence judgments through a different mechanism than consensus independence. Together, these results indicate that the cognitive processes underlying sensitivity to source dependency may differ from those underlying sensitivity to consensus independence.

```{r}
#| label: tbl-m-comp-e3-rm-ind
#| tbl-cap: "Experiment 2 Model comparison using LOOIC (lower LOOIC = better performance), comparing Dependent vs. Dependent Source (excluding Independent)"
load(here("analyses/output/looic-model-comparison-e2-rm-independent.Rdata"))
modelComparisonTable(model_looic_rm_independent, model_full_names)
```

```{r}
#| label: tbl-params-e2-rm-ind
#| tbl-cap: "Parameter estimates of the best performing model in Experiment 2 comparing Dependent vs. Dependent Source: the *Start* model, in which the Start parameter is the only parameter that can vary across consensus conditions."

load(here("analyses/output/fe-brms-exp-2-power-fit-start-rm-independent.Rdata"))
param_names <- c(
                "Start_{Dependent}",
                "Start_{Dependent Source}",
                "Increase",
                 "Rate")
extractEstimates(fit, param_names)
```

# References

-   Bond, R. (2005). Group Size and Conformity. *Group Processes & Intergroup Relations*, *8*(4), 331–354. <https://doi.org/10.1177/1368430205056464>

-   Latané, B. (1981). The psychology of social impact. *American Psychologist*, *36*(4), 343–356. <https://doi.org/10.1037/0003-066X.36.4.343>

-   Paul-Christian Buerkner (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. *Journal of Statistical Software*, 80(1), 1-28. `doi:10.18637/jss.v080.i01`

-   Paul-Christian Buerkner (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. *The R Journal*. 10(1), 395–411. `doi:10.32614/RJ-2018-017`

-   The Stan Development Team (n.d). *Stan Modeling Language User's Guide and Reference Manual*. <https://mc-stan.org/users/documentation/>.

-   Stan Development Team (2020). RStan: the R interface to Stan. R package version 2.21.2. <https://mc-stan.org/>

-   Vehtari, A., Gelman, A., & Gabry, J. (2017). Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC. *Statistics and Computing*, *27*(5), 1413–1432. <https://doi.org/10.1007/s11222-016-9696-4>

# Supplementary Materials

I have just thrown in a few extra analyses/information that might be useful to include in the supplementary materials (you might also think they belong in the main manuscript, too). They haven't all been linked in text though.

# Comparing prior belief vs. baseline

A relevant question for relevant for our modelling approach is how to deal with participants' reported prior beliefs compared to the baseline condition, where there is 1 person agreeing with the claim, and 1 person disagreeing. In theory, it would be nice to control for people's prior beliefs in some way. In all of my previous consensus studies we have done this, in order to account for the fact that people who already have strong prior beliefs are less likely to shift them. There are a few problems with this in the current study, though.

1.  There is only one trial per participant, so you cannot control for individual variation in priors in the same way as you normally would (i.e., with a random intercept).
2.  As I will show, the prior and baseline condition are *highly* correlated, meaning that controlling for controlling for prior belief as a fixed effect could be problematic, and artificlally reduce the predictive power of the model and result in issues of multicollinearity.

## Does confidence change from prior to baseline?

It's worth seeing whether confidence at prior is the same as at baseline. This is tricky because they were measured on different scales, but we can see if they're correlated at least. 

```{r}
experiments = c("1-combined", "2")

source(here("analyses/prior-baseline-comparison.R"))

all_corr_plots <- vector("list", length = length(experiments))

for (i in 1:length(experiments)){
  exp <- experiments[i]
  all_corr_plots[[i]] <- checkPriorBaselineCorr(exp)
}


combined_corr_plot <- ggarrange(plotlist = all_corr_plots)
combined_corr_plot
```

These plots show very high correlations between someone's prior belief and their confidence at baseline in E2, sugesting they correspond pretty well with each other. 

## Models

In this section, we show the precise formalization of the different models that we ran in our model comparison, as they were implemented in BRMS. Note that in our models, all parameters were exponentiated so that they are constrained to be positive. Full model code can be found at \[link to model code\].

All models assumed a power function relationship between consensus size and confidence, as follows:

```         
  confidence ~ (exp(start) + exp(increase)) - exp(increase) * (n_sources^(-exp(rate)))
```

What changed between models was how the fixed effects were defined. In other words, does the model assume the parameters vary as a function consensus condition?

### *None* model: assumes neither *start, increase,* nor *rate* vary as a function of consensus

```         
      start ~ 1,
      increase ~ 1, 
      rate ~ 1, 
```

#### *Start* model: assumes that only the *start* parameter varies as a function of consensus.

```         
 start ~ 1 + consensus
 increase ~ 1
  rate ~ 1
```

#### *Increase* model: assumes that only the *increase* parameter varies as a function of consensus.

```         
start ~ 1
increase ~ 1 + consensus
rate ~ 1
```

#### *Rate* model: assumes that only the *rate* parameter varies as a function of consensus.

```         
start ~ 1
increase ~ 1
rate ~ 1 + consensus
```

#### *Start + Increase* model: assumes that both the *start* and *increase* parameters vary as a function of consensus.

```         
start ~ 1 + consensus
increase ~ 1 + consensus
rate ~ 1
```

#### *Start + Rate* model: assumes that both the *start* and *rate* parameters vary as a function of consensus.

```         
start ~ 1 + consensus
increase ~ 1
rate ~ 1 + consensus
```

#### *Increase + Rate* model: assumes that both the *increase* and *rate* parameters vary as a function of consensus.

```         
start ~ 1
increase ~ 1 + consensus
rate ~ 1 + consensus
```

#### *Full* model: assuming *start, increase,* and *rate* all vary as a function of consensus condition.

```         
      start ~ 1 + consensus,
      increase ~ 1 + consensus,
      rate ~ 1 + consensus
```

#### 

### Priors

We placed weakly informative priors on all parameters. Intercepts were given log-normal priors reflecting plausible values, while consensus effects (slopes) were centered at zero with moderate variance:

-   **Intercepts**

$$
 Start \sim \mathcal{N}(\log 50, 0.3) 
$$ $$
  Increase \sim \mathcal{N}(\log 25, 0.3) 
$$

$$
  Rate \sim \mathcal{N}(\log 0.5, 0.15) 
$$

-   **Condition Effects (when included)**\
    Start and Increase:

$$
\mathcal{N}(0, 0.2) 
$$

Rate

$$
\mathcal{N}(0, 0.15) 
$$
