---
title: "Modelling the impact of increasing sources using Bayesian, non-linear (power function) regressions"
author: "Manikya Alister"
format: 
  html:
   embed-resources: true
editor: visual
message: false
warning: false
code-fold: true
---

```{r}
library(tidyverse)
library(posterior)
library(here)
library(ggdist)
library(raincloudplots)


# define experiments
experiments = c(3)

# load model fits
fits = list()
for (exp in experiments){
load(here(paste0("output/fe-brms-exp",exp,"-power-fit-full.Rdata")))
fits[[exp]] <- fit
}


```

There are a few reasons why I think modelling this kind of problem using non-linear regression, especially one that follows a power (or any asymptotic growth) function is useful. First, it has unique predictive value, allowing us to estimate how increasing the number of sources will increase confidence, even if we do not have direct empirical measurements for a given source number. This can therefore allow us to do things like predict when adding new sources will stop increasing confidence as a function of different types of consensus manipulations (e.g., independence), with minimal data.

This type of modelling also corresponds well to theoretical models describing the relationship between persuasion and social group size (Bond, 2005).

Third, if there are differences in how independence (or any other manipulation, like expertise) affects the consensus X source number relationship, non-linear modelling like this allows us to make inferences about what is driving these differences. Using a power function, like I do here, we can infer whether source independence affects the rate at which confidence increases with multiple sources, whether it affects the highest possible confidence (asymptote) that increasing the number of sources will produce, or both. Therefore, I think doing this kind of analysis can provide added benefit beyond something like a mixed source number X independence ANOVA. That being said, you might not think the extra insights that can be gained by this approach are that important for the current study, which I would completely understand! I have been wanting to try this kind of analysis for a while, so it won't have been a waste of time for me if we don't end up using it.

# Comparing prior belief vs. baseline

A relevant question for relevant for our modelling approach is how to deal with participants' reported prior beliefs compared to the baseline condition, where there is 1 person agreeing with the claim, and 1 person disagreeing. In theory, it would be nice to control for people's prior beliefs in some way. In all of my previous consensus studies we have done this, in order to account for the fact that people who already have strong prior beliefs are less likely to shift them. There are a few problems with this in the current study, though.

1.  There is only one trial per participant, so you cannot control for individual variation in priors in the same way as you normally would (i.e., with a random intercept).
2.  As I will show, the prior and baseline condition are *highly* correlated, meaning that controlling for controlling for prior belief as a fixed effect could be problematic, and artificlally reduce the predictive power of the model and result in issues of multicollinearity.

## Does confidence change from prior to baseline?

If we can treat prior and baseline as essentially the same, this will vastly simplify our modelling, as we can treat the baseline condition as people's priors. First, let's check the correlation

```{r}
source(here("analyses/prior-baseline-comparison.R"))

all_corr_plots <- vector("list", length = length(experiments))

for (i in 1:length(experiments)){
  all_corr_plots[[i]] <- checkPriorBaselineCorr(i)
}


combined_corr_plot <- ggarrange(plotlist = all_corr_plots)
combined_corr_plot
```

These plots show very high correlations between someone's prior belief and their confidence at baseline. Now lets compare the means of the groups and their distributions.

```{r}

all_group_plots <- vector("list", length = length(experiments))

for (i in 1:length(experiments)){
  exp <- experiments[i]
  all_group_plots[[i]] <- checkPriorBaselineGroups(exp)
}


combined_group_plot <- ggarrange(plotlist = all_group_plots)
combined_group_plot
```

These plots show that scores are pretty much the same from prior to baseline. Although some participants do seem to change their rating, this appears to be the vast minority, and at the group level the distributions look similar. The means are also not significantly different from each other.

Something else that's interesting is that the baseline condition seems to bring the distribution more towards the middle, somewhat nullifying the effect of priors (i.e., the baseline distribution is more symmetrical around 0).

Taken together, I would argue that it's safe to ignore prior beliefs in the modelling and instead focus only on how beliefs change from baseline.

# Evaluating model fit

Here, we can see whether the functional (power) form, describes the relationship between confidence and number of sources for each dependence condition.

```{r}
plotFit(exp = 1,
        title = "Experiment 1",
        subtitle = NULL)
```

```{r}
plotFit(exp = 2,
        title = "Experiment 2",
        subtitle = NULL)
```

```{r}
plotFit(exp = 3,
        title = "Experiment 3",
        subtitle = NULL)
```

Seems like it does pretty well! I've also had a look at some other similar functions, like log or exponential functions, and found that log & power fit similarly better than exponential, but I chose power because I find the parameters more interperetable.

Now that we know the model fits well, we can investigate what is driving any differences between conditions. Specifically, whether they are driven by differences in rate of increase, asymptote, or both.

# Model comparison

Another way to try and figure out which parameters are varying across conditions is to perform a model comparison, where we compare models that either allow certain parameters to vary across consensus conditions or not. In other words, it lets us figure out whether assuming that a parameter, such as rate, is different in the independent vs. dependent vs. dependent source condition actually makes the fit relative to complexity better. I ran 8 models, with all possible combination of parameters varying as a function of consensus, where "full" is all parameters varying, and "none" is assuming that parameters are the same across consensus conditions.

```{r}
load(file = here("output/looic-model-comparison.Rdata"))
print("Results of model comparison ordered from best (left) to worst (right) according to LOOIC: ")
model_looic 
```

The results suggest that only allowing the increase (asymptote) parameter to vary across time results in the best model performance relative to complexity. It's worth noting though, that the difference between all of the models that assume at least one parameter varies across consensus conditions is quite small (i.e., all models except "none" are separated by only a few points). This fairly large difference between full vs. none suggests so me that there is pretty strong differences across consensus conditions, but this could be mainly driven by the independent condition being quite different from the dependent and dependent source conditions.

Overall, between the model comparison and the parameter estimates I'd interpret this as there being some evidence there is a) differences between conditions and that b) these difference are driven by asymptote. asymptote between con between independent and the rest, and some evidence to suggest differences beteween dependent and dependent_source, but they're smaller and less reliable.

# Compare parameters across conditions

The table below shows the absolute means and 95% credible differences between parameters, as well as the *relative* means and 95% credible differences between parameters across independence conditions. The relative differences were calculated because the parameters are on different scales (small differences in rate result in big differences in shape of the curve), so this allows us to see by what % is the parameter in the independent condition than the equivalent in the dependent condition.

```{r}
param_contrasts_summary <- NULL
param_contrasts_long <- NULL
intercept_level <- c("Dependent" = "consensusdependent")
comparison_levels <- c("Independent" = "consensusindependent", "Dependent Source" = "consensusdependent_source")


for (exp in experiments) {
  fit <- fits[[exp]]
  
  print("Model fit: ")  
  #print(summary(fit)$fixed %>% select(-Bulk_ESS,-Tail_ESS))
  
  # Extract draws
  draws <- as_draws_df(fit)
  
  # 2. Helper to extract and transform parameters
  get_param_contrasts <- function(draws,
                                  param_name,
                                  condition_effect_name) {
    # asymptote is start + increase
    if (param_name == "asymptote") {
      start_intercept <- draws[["b_start_Intercept"]]
      increase_intercept <- draws[["b_increase_Intercept"]]
      
      intercept <- exp(start_intercept) + exp(increase_intercept)
      condition_effect <- exp(start_intercept + draws[[paste0("b_start_", condition_effect_name)]]) + exp(increase_intercept + draws[[paste0("b_increase_", condition_effect_name)]]) #+ intercept
      
    } else{
      intercept_raw <- draws[[paste0("b_", param_name, "_Intercept")]]
      intercept <- exp(intercept_raw)
      condition_effect <- exp(intercept_raw + draws[[paste0("b_", param_name, "_", condition_effect_name)]])
    }
    
    tibble(
      parameter = param_name,
      condition_effect_name = condition_effect_name,
      condition_effect = condition_effect,
      intercept = intercept,
      abs_diff = intercept - condition_effect,
      rel_diff = (intercept - condition_effect) / condition_effect
    )
  }
  
  # Individual parameter contrasts
  log_params <- c("rate", "asymptote", "start")
  
  for (i in 1:length(comparison_levels)) {
    comparison_level <- comparison_levels[i]
    param_contrasts <- bind_rows(lapply(log_params, function(p)
      get_param_contrasts(draws, p, comparison_level)))
    
    # 5. Summarise posterior contrasts
    param_contrast_summ_exp <- param_contrasts %>%
      group_by(parameter) %>%
      summarise(
        mean_cond_effect = mean(condition_effect),
        mean_intercept = mean(intercept),
        mean_abs_diff = mean(abs_diff),
        ci_abs_lower = quantile(abs_diff, 0.025),
        ci_abs_upper = quantile(abs_diff, 0.975),
        prob_intercept_gt_contrast = mean(abs_diff > 0),
        prob_intercept_lt_contrast = mean(abs_diff < 0),
        prob_intercept_diff_contrast = prob_intercept_gt_contrast - prob_intercept_lt_contrast,
        
        
        mean_rel_diff = mean(rel_diff),
        ci_rel_lower = quantile(rel_diff, 0.025),
        ci_rel_upper = quantile(rel_diff, 0.975)
      ) %>%
      arrange(match(parameter, c(
        "rate", "increase", "start", "asymptote"
      )))
    
    print(paste0(names(intercept_level), " vs. ", names(comparison_level)))
    #param_contrast_summ_exp <- param_contrasts_summary[[1]]
    
    print(knitr::kable(
      param_contrast_summ_exp %>% select(
        parameter,
        mean_cond_effect,
        mean_intercept,
        mean_abs_diff,
        ci_abs_lower,
        ci_abs_upper,
        mean_rel_diff,
        ci_rel_lower,
        ci_rel_upper
      )
    ))

  
    print(
      paste0(
        "Probability that parameter in ",
        names(intercept_level),
        " condition > ",
        names(comparison_level),
        " condition:"
      )
    )
    print(param_contrast_summ_exp %>% select(parameter, prob_intercept_gt_contrast))
    print(
      paste0(
        "Probability that parameter in ",
        names(intercept_level),
        " condition < ",
        names(comparison_level),
        " condition:"
      )
    )
    print(param_contrast_summ_exp %>% select(parameter, prob_intercept_lt_contrast))
    
    param_contrasts_summary[[exp]] <- param_contrast_summ_exp %>%
      filter(parameter != "increase")
    
    param_contrasts_long_exp <- param_contrasts %>%
      pivot_longer(
        cols = c(condition_effect, intercept),
        names_to = "consensus",
        values_to = "value"
      ) %>%
      mutate(
        experiment = exp,
        comparison = case_when(
          condition_effect_name == "consensusindependent" ~ "Dependent vs. Independent",
          condition_effect_name == "consensusdependent_source" ~ "Dependent vs. Dependent Source"
        )
      )
    
    param_contrasts_long <- bind_rows(param_contrasts_long, param_contrasts_long_exp)
  }
}
```

It looks like that for neither the independent vs. dependent comparison, nor the dependent vs dependent_source comparison, are any parameters fully reliably different at a 95% credible interval (there's some overlap). But it does seem still much closer for the independent comparison than the dependent_source comparison. Also, when you look at how probable it is that asymptote in the Dependent condition \< Independent condition, there is a 97% probability that there is a difference. FOr dependent vs. dependent_source, there's an 87% probability that dependnent \> dependent_source. So even though they're small differences, and the 95% credible interval are not showing completely reliable differences, there is still some evidence to suggest differences, especially in the independent comparison-- it just depends on how conservative you want to be. The plot show this more clearly.

```{r}
# Plot

param_contrasts_long %>%
  filter(parameter != "increase") %>% # remove increase parameter because it is basically the same as asymptote
ggplot(aes(x = value, colour = consensus, fill = consensus)) +
  stat_halfeye(alpha = 0.6, slab_linewidth = 0.5, slab_color = "black",  normalize = "xy", .width = .95) +
  #facet_wrap(condition_effect_name~parameter, scales = "free_x")+
  facet_grid(comparison ~ parameter, scales = "free") +
  theme_minimal() +
  labs(
    title = "Posterior Distributions by Consensus Type and Experiment",
   # subtitle = paste0("Experiment ", experiment),
    x = "Estimate",
    y = NULL,
    fill = "Consensus",
    color = "Consensus"
  )
```

Above, the plot shows that there are much more reliable differences between the independent and dependent asymptote compared to the dependent vs. dependent source. Something that could be messing the asymptote reliability up a bit though is the difference in the starting points. For some reason, the samples in the dependent source group had a lower starting point in the baseline condition, which is a bit strange, but could be suppressing the influence of asymptote.

One thing that I've done is just look at the probability that a a paramter in one condition is greater than that of the other condition by looking at the proportion of samples where the difference between conditions was greater than 0. Looking at this, it seems like much stronger evidence for asymptote differences across conditions, with 97% probability that asymptote in the independent condition is greater than that in dependent, and 88% probability that asymptote in the dependent condition is greater than that in the dependent source condition. When we look at the same for rate, these probabilities are close to 50%, suggesting that any differences are most likely being driven by asymptote.
