---
title: "Evaluating Non-Linear Modelling"
format: html
editor: visual
---

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

# Which functional form fits the data best?

I tested an exponential function, power function, and logarithmic function.

## Experiment 1

```{r}
extractLooic <- function(model_list, condition = NULL) {
  filtered_list <- if (!is.null(condition)) {
    model_list[sapply(model_list, function(x) x$cond == condition)]
  } else {
    model_list
  }
  
  results <- data.frame(
    #model = names(filtered_list),
    func = sapply(filtered_list, function(x) x$func),
    looic = sapply(filtered_list, function(x) x$looic),
    looic_se = sapply(filtered_list, function(x) x$looic_se)
  ) %>%
    mutate(
      rank = rank(looic)
    )
  
  results <- arrange(results, looic)
  
  results
}
load(here("output/all-output-e1.Rdata"))
extractLooic(all_output, condition = "independent")
extractLooic(all_output, condition = "dependent")

```

## Experiment 2

```{r}
load(here("output/all-output-e2.Rdata"))
extractLooic(all_output, condition = "independent")
extractLooic(all_output, condition = "dependent")
```

For both the independent and dependent condition, across both experiments, the **exponential** function seemed to fit best, suggesting that the affect of additional sources on confidence in a consensus may follow an exponential form.

```{r}
# plotting
experiments <- c(1,2)
plotting_data <- NULL

extractParameters = function(all_output, condition, model){
  model_cond <- paste0(condition,"+",model)
  # get function
    #func = all_output[[model_cond]]$func
    # get model parameter values
    params = all_output[[model_cond]]$estimates
    params
}

# function for simulating the exponential function
simulateExp = function(x, stat, n_sources = 0:11){
  a <- x["a",stat]
  b <- x["a",stat]
  c <- x["c",stat]
  
 a + b * (1 - exp(-c * n_sources))

}

d_plotting_human <- NULL
d_plotting_model <- NULL
conditions <- c("independent", "dependent")

for (i in experiments){
  for (cond in conditions) {
       print(cond)
  # load output
  load(here(paste0("output/all-output-e",i,".Rdata")))
  n_sources <- 0:11
  params <- extractParameters(all_output, cond, "exp")
  
  sim_data <- data.frame(
    model_predictions = simulateExp(params, "50%", n_sources),
    upper_CI = simulateExp(params, "97.5%", n_sources),
    lower_CI = simulateExp(params, "2.5%", n_sources),
    n_sources = n_sources,
    consensus = cond,
    experiment = i
  )
  
  d_plotting_model <- rbind(d_plotting_model, sim_data)
  
  # load real data
  load(here(paste0("data/experiment-",i,"/clean/d-n-repetitions.Rdata")))

  # summarise by repetition
  d_summ <- data %>%
  group_by(consensus, n_sources) %>%
  summarise(median = median(confidence), sd = sd(confidence), n = n())%>%
    mutate(se = sd/sqrt(n),
    experiment = i)
  
  d_plotting_human <- rbind(d_plotting_human, d_summ)
  
}}

d_plotting_model %>% 
  ggplot(aes(
    x = n_sources,
    y = model_predictions,
    group = consensus,
    colour = consensus
  )) +
  geom_line() +
  geom_point(data = d_plotting_human, aes(y = median))+
  facet_wrap(~experiment)

```
